---
title: "Programming Assignment 3"
author: "Okky Mabruri"
date: "13/09/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/mnt/sdb2/OneDrive/Who-Am-I/datasci-ai/course-data-science-specialization-JHU/2-R-Programming")
```

credit : https://rpubs.com/SLamara


# Introduction

The data for this assignment come from the Hospital Compare web site [http://hospitalcompare.hhs.gov](http://hospitalcompare.hhs.gov) run by the U.S. Department of Health and Human Services. The purpose of the web site is to provide data and information about the quality of care at over 4,000 Medicare-certified hospitals in the U.S. This dataset essentially covers all major U.S. hospitals.  This dataset is used for a variety of purposes, including determining whether hospitals should be fined for not providing high quality care to patients (see [http://goo.gl/jAXFX](http://goo.gl/jAXFX) for some background on this particular topic).


The  Hospital  Compare  web  site  contains  a  lot  of  data  and  we  will  only  look  at  a  small  subset  for  this assignment.  The zip file for this assignment contains three files

* outcome-of-care-measures.csv: Contains information about 30-day mortality and readmission rates for heart attacks, heart failure, and pneumonia for over 4,000 hospitals.

* hospital-data.csv:  Contains information about each hospital.

* Hospital_Revised_Flat files.pdf:  Descriptions of the variables in each file (i.e the code book).


# 1  Plot the 30-day mortality rates for heart attack

Read the outcome data into R via theread.csv


``` {r}
setwd('ProgAssignment3-data/')
hospitaldata <- read.csv('hospital-data.csv')
outcome <-
  read.csv("outcome-of-care-measures.csv", colClasses = "character")
```

## 1.1 Check outcome data

```{r}
nrow(outcome)
```
```{r}
names(outcome[1:11])
```

## 1.2 Make hist plot
```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11])
```

# 2 Finding the best hospital in a state

Handling ties.  If there is a tie for the best hospital for a given outcome, then the hospital names should be sorted in alphabetical order and the first hospital in that set should be chosen (i.e.  if hospitals “b”, “c”,and “f” are tied for best, then hospital “b” should be returned).The function should use the following template.

```{R}
best <- function(state, outcome) {
  
  ## Read outcome data
  setwd('ProgAssignment3-data/')
  
  data <-
    read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that state and outcome are valid
  
  if (!(state %in% data$State)) {
    result <- "invalid state"
  }
  else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
    result <- "invalid outcome"
  }
  else{
    keys <-
      c(
        "heart attack" = 11,
        "heart failure" = 17,
        "pneumonia" = 23
      )
  
    outcomeKey <- keys[outcome]
    
    ## Return hospital name in that state with lowest 30-day death rate
    
    dataPerState <- split(data, data$State)
    dataOurState <- dataPerState[[state]]
    dataOurState <-
      dataOurState[order(dataOurState["Hospital.Name"]),]
    dataOutcome <-
      suppressWarnings(as.numeric(dataOurState[, outcomeKey]))
    good <- complete.cases(dataOutcome)
    dataOutcome <- dataOutcome[good]
    dataOurState <- dataOurState[good, ]
    minimum <- min(dataOutcome)
    index <- match(minimum, dataOutcome)
    result <- dataOurState[index, 2]
  }
  result
}
```

```{R}
best("TX", "heart attack")
```

```{R}
best("TX", "heart failure")
```

```{R}
best("MD", "heart attack")
```

```{R}
best("MD", "pneumonia")
```

```{R}
best("BB", "heart attack")
```

```{R}
best("NY", "hert attack")
```
# 3  Ranking hospitals by outcome in a state

Write a function called rank hospital that takes three arguments:  the 2-character abbreviated name of a state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num).The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the ranking specified by the num argument.  For example, the call 

```
rankhospital("MD", "heart failure", 5)
```

would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate for heart failure.  The num argument can take values “best”, “worst”,  or an integer indicating the ranking(smaller numbers are better).  If the number given by num is larger than the number of hospitals in that state, then the function should return NA. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.

Handling ties.  It may occur that multiple hospitals have the same 30-day mortality rate for a given causeof death.  In those cases ties should be broken by using the hospital name.  

Note that Cypress Fairbanks Medical Center and Detar Hospital Navarro both have the same 30-day rate(8.7).   However,  because  Cypress  comes  before  Detar  alphabetically,  Cypress  is  ranked  number  3  in  this scheme  and  Detar  is  ranked  number  4.   One  can  use  the order function  to  sort  multiple  vectors  in  this manner (i.e.  where one vector is used to break ties in another vector).

```{r}
rankhospital <- function(state, outcome, num = "best") {
  ## Read outcome data
  
  setwd('ProgAssignment3-data/')
  
  data <-
    read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that state and outcome are valid
  
  if (!(state %in% data$State)) {
    result <- "invalid state"
  }
  else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
    result <- "invalid outcome"
  }
  else {
    keys <-
      c(
        "heart attack" = 11,
        "heart failure" = 17,
        "pneumonia" = 23
      )
    outcomeKey <- keys[outcome]
    
    
    ## Return hospital name in that state with the given rank
    ## 30-day death rate
    
    dataPerState <- split(data, data$State)
    dataOurState <- dataPerState[[state]]
    dataOutcome <-
      suppressWarnings(as.numeric(dataOurState[, outcomeKey]))
    good <- complete.cases(dataOutcome)
    dataOutcome <- dataOutcome[good]
    dataOurState <- dataOurState[good, ]
    dataOurState <-
      dataOurState[order(dataOutcome, dataOurState["Hospital.Name"]), ]
    if (grepl("^[0-9]+$", num)) {
      if (as.numeric(num) > length(dataOutcome)) {
        result <- NA
      }
      else {
        result <- dataOurState[as.numeric(num), "Hospital.Name"]
      }
    }
    else if (num == "best") {
      result <- dataOurState[1, "Hospital.Name"]
    }
    else if (num == "worst") {
      result <- dataOurState[length(dataOutcome), "Hospital.Name"]
    }
    else
      result <- NA
  }
  result
}
```

```{R}
rankhospital("TX", "heart failure", 4)
```
```{R}
rankhospital("MD", "heart attack", "worst")
```
```{R}
rankhospital("MN", "heart attack", 5000)
```

# 4 Ranking hospitals in all states

I implement a function rankAll which takes as arguments the outcome name (outcome) and hospital ranking (num) and returns a 2-column data frame containing the hospital in each state that has the ranking specified in num.

The function returns a value for every state (some may be NA). The first column in the data frame contains the hospital name and the second one contains the 2-character abbreviation for the state name. Hospitals that do not have data on a particular outcome are excluded from the set of hospitals when deciding the rankings.

Although it is possible to call the rankHospital function from the previous section, I decided, for didactic purposes, not using it.

```{R}
rankall <- function(outcome, num = "best") {
  setwd('ProgAssignment3-data/')
  
  dataAll <-
    data.frame(hospital = character(), state = character())
  
  ## Read outcome data
  
  data <-
    read.csv("outcome-of-care-measures.csv", colClasses = "character")
  
  ## Check that outcome and num are valid
  
  if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
    dataAll <- "invalid outcome"
  }
  else {
    keys <-
      c(
        "heart attack" = 11,
        "heart failure" = 17,
        "pneumonia" = 23
      )
    outcomeKey <- keys[outcome]
    
    ## For each state, find the hospital of the given rank
    
    dataPerState <- split(data, data$State)
    for (stat in names(dataPerState)) {
      dataOurState <- dataPerState[[stat]]
      dataOutcome <-
        suppressWarnings(as.numeric(dataOurState[, outcomeKey]))
      good <- complete.cases(dataOutcome)
      dataOutcome <- dataOutcome[good]
      dataOurState <- dataOurState[good, ]
      dataOurState <-
        dataOurState[order(dataOutcome, dataOurState["Hospital.Name"]),]
      
      if (num == "best") {
        numState <- c(1)
      } else {
        if (num == "worst") {
          numState <- length(dataOutcome)
        } else {
          numState <- num
        }
      }
      
      dataPart <-
        data.frame(hospital = dataOurState[numState, "Hospital.Name"],
                   state = stat,
                   row.names = stat)
      dataAll <- rbind(dataAll, dataPart)
    }
  }
  
  ## Return a data frame with the hospital names and the (abbreviated) state name
  
  dataAll
}
```

## Testing rankAll

```{R}
head(rankall("heart attack", 20), 10)
```

```{R}
tail(rankall("pneumonia", "worst"), 3)
```

```{R}
tail(rankall("heart failure"), 10)
```
# Programming Assignment 3: Quiz

```{R}
best("SC", "heart attack")
```
```{R}
best("NY", "pneumonia")
```
```{R}
best("AK", "pneumonia")
```
```{R}
rankhospital("NC", "heart attack", "worst")
```
```{R}
rankhospital("WA", "heart attack", 7)
```
```{R}
rankhospital("TX", "pneumonia", 10)
```
```{R}
rankhospital("NY", "heart attack", 7)
```
```{R}
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)
```
```{R}
r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```
```{R}
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)
```